---
title: "IMDB_RNN"
author: "Dolores Chu"
date: "3/9/2020"
output: html_document
---

Modifying the network.
```{r}
library(keras)
embedding_layer <- layer_embedding(input_dim = 1000, output_dim = 64) 
```

```{r}
# Consider the top 10,000 words
max_features <- 10000
# Cut texts after 150 words
maxlen <- 150
# Load the data as lists of integers.
imdb <- dataset_imdb(num_words = max_features)
c(c(x_train, y_train), c(x_test, y_test)) %<-% imdb
x_train <- pad_sequences(x_train, maxlen = maxlen)
x_test <- pad_sequences(x_test, maxlen = maxlen)
```

Modifying taining and validation dataset sizes
```{r}
indices <- sample(1:10100, size = 100)
x_training <- x_train[indices, ]
y_training <- y_train[indices]
x_validation <- x_train[-indices, ]
y_validation <- y_train[-indices]
```



```{r}
model <- keras_model_sequential() %>% 
  # We specify the maximum input length to our Embedding layer
  # so we can later flatten the embedded inputs
  layer_embedding(input_dim = 10000, output_dim = 8, 
                  input_length = maxlen) %>% 
  # We flatten the 3D tensor of embeddings 
  # into a 2D tensor of shape `(samples, maxlen * 8)`
  layer_flatten() %>% 
  # We add the classifier on top
  layer_dense(units = 1, activation = "sigmoid") 
model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_training, y_training,
  epochs = 10,
  batch_size = 32, validation_data = list (x_validation, y_validation))
```

Using pre-trained word embeddings.
```{r}
imdb_dir <- "/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/aclImdb"
train_dir <- file.path(imdb_dir, "train")
labels <- c()
texts <- c()
for (label_type in c("neg", "pos")) {
  label <- switch(label_type, neg = 0, pos = 1)
  dir_name <- file.path(train_dir, label_type)
  for (fname in list.files(dir_name, pattern = glob2rx("*.txt"), 
                           full.names = TRUE)) {
    texts <- c(texts, readChar(fname, file.info(fname)$size))
    labels <- c(labels, label)
  }
}

library(keras)
maxlen <- 150                 
training_samples <- 100      
validation_samples <- 10000  
max_words <- 10000            
tokenizer <- text_tokenizer(num_words = max_words) %>% 
  fit_text_tokenizer(texts)
sequences <- texts_to_sequences(tokenizer, texts)
word_index = tokenizer$word_index
cat("Found", length(word_index), "unique tokens.\n")

data <- pad_sequences(sequences, maxlen = maxlen)
labels <- as.array(labels)
cat("Shape of data tensor:", dim(data), "\n")

cat('Shape of label tensor:', dim(labels), "\n")

indices <- sample(1:nrow(data))
training_indices <- indices[1:training_samples]
validation_indices <- indices[(training_samples + 1): 
                              (training_samples + validation_samples)]
x_train <- data[training_indices,]
y_train <- labels[training_indices]
x_val <- data[validation_indices,]
y_val <- labels[validation_indices]

glove_dir = '/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/glove.6B'
lines <- readLines(file.path(glove_dir, "glove.6B.100d.txt"))
embeddings_index <- new.env(hash = TRUE, parent = emptyenv())
for (i in 1:length(lines)) {
  line <- lines[[i]]
  values <- strsplit(line, " ")[[1]]
  word <- values[[1]]
  embeddings_index[[word]] <- as.double(values[-1])
}

embedding_dim <- 100
embedding_matrix <- array(0, c(max_words, embedding_dim))
for (word in names(word_index)) {
  index <- word_index[[word]]
  if (index < max_words) {
    embedding_vector <- embeddings_index[[word]]
    if (!is.null(embedding_vector))
      # Words not found in the embedding index will be all zeros.
      embedding_matrix[index+1,] <- embedding_vector
  }
}

model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
summary(model)

get_layer(model, index = 1) %>% 
  set_weights(list(embedding_matrix)) %>% 
  freeze_weights()

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
save_model_weights_hdf5(model, "pre_trained_glove_model.h5")
plot(history)
```

Task-specific embedding
```{r}
model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
plot(history)
```
When the training sample is as small as 100, pre-trained word embeddings outperformed task-specific embedding, which is only doing as good of a job as baseline (about 0.5 accuracy).

Now increase the training sample size to 300.
Pre-trained word embeddings.
```{r}
imdb_dir <- "/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/aclImdb"
train_dir <- file.path(imdb_dir, "train")
labels <- c()
texts <- c()
for (label_type in c("neg", "pos")) {
  label <- switch(label_type, neg = 0, pos = 1)
  dir_name <- file.path(train_dir, label_type)
  for (fname in list.files(dir_name, pattern = glob2rx("*.txt"), 
                           full.names = TRUE)) {
    texts <- c(texts, readChar(fname, file.info(fname)$size))
    labels <- c(labels, label)
  }
}

library(keras)
maxlen <- 150                 
training_samples <- 300      
validation_samples <- 10000  
max_words <- 10000            
tokenizer <- text_tokenizer(num_words = max_words) %>% 
  fit_text_tokenizer(texts)
sequences <- texts_to_sequences(tokenizer, texts)
word_index = tokenizer$word_index
cat("Found", length(word_index), "unique tokens.\n")

data <- pad_sequences(sequences, maxlen = maxlen)
labels <- as.array(labels)
cat("Shape of data tensor:", dim(data), "\n")

cat('Shape of label tensor:', dim(labels), "\n")

indices <- sample(1:nrow(data))
training_indices <- indices[1:training_samples]
validation_indices <- indices[(training_samples + 1): 
                              (training_samples + validation_samples)]
x_train <- data[training_indices,]
y_train <- labels[training_indices]
x_val <- data[validation_indices,]
y_val <- labels[validation_indices]

glove_dir = '/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/glove.6B'
lines <- readLines(file.path(glove_dir, "glove.6B.100d.txt"))
embeddings_index <- new.env(hash = TRUE, parent = emptyenv())
for (i in 1:length(lines)) {
  line <- lines[[i]]
  values <- strsplit(line, " ")[[1]]
  word <- values[[1]]
  embeddings_index[[word]] <- as.double(values[-1])
}

embedding_dim <- 100
embedding_matrix <- array(0, c(max_words, embedding_dim))
for (word in names(word_index)) {
  index <- word_index[[word]]
  if (index < max_words) {
    embedding_vector <- embeddings_index[[word]]
    if (!is.null(embedding_vector))
      # Words not found in the embedding index will be all zeros.
      embedding_matrix[index+1,] <- embedding_vector
  }
}

model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
summary(model)

get_layer(model, index = 1) %>% 
  set_weights(list(embedding_matrix)) %>% 
  freeze_weights()

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
save_model_weights_hdf5(model, "pre_trained_glove_model.h5")
plot(history)
```

Now compare with task-specific embedding
```{r}
model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
plot(history)
```

Now increase the training sample size to 600.
Pre-trained word embeddings.
```{r}
imdb_dir <- "/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/aclImdb"
train_dir <- file.path(imdb_dir, "train")
labels <- c()
texts <- c()
for (label_type in c("neg", "pos")) {
  label <- switch(label_type, neg = 0, pos = 1)
  dir_name <- file.path(train_dir, label_type)
  for (fname in list.files(dir_name, pattern = glob2rx("*.txt"), 
                           full.names = TRUE)) {
    texts <- c(texts, readChar(fname, file.info(fname)$size))
    labels <- c(labels, label)
  }
}

library(keras)
maxlen <- 150                 
training_samples <- 600      
validation_samples <- 10000  
max_words <- 10000            
tokenizer <- text_tokenizer(num_words = max_words) %>% 
  fit_text_tokenizer(texts)
sequences <- texts_to_sequences(tokenizer, texts)
word_index = tokenizer$word_index
cat("Found", length(word_index), "unique tokens.\n")

data <- pad_sequences(sequences, maxlen = maxlen)
labels <- as.array(labels)
cat("Shape of data tensor:", dim(data), "\n")

cat('Shape of label tensor:', dim(labels), "\n")

indices <- sample(1:nrow(data))
training_indices <- indices[1:training_samples]
validation_indices <- indices[(training_samples + 1): 
                              (training_samples + validation_samples)]
x_train <- data[training_indices,]
y_train <- labels[training_indices]
x_val <- data[validation_indices,]
y_val <- labels[validation_indices]

glove_dir = '/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/glove.6B'
lines <- readLines(file.path(glove_dir, "glove.6B.100d.txt"))
embeddings_index <- new.env(hash = TRUE, parent = emptyenv())
for (i in 1:length(lines)) {
  line <- lines[[i]]
  values <- strsplit(line, " ")[[1]]
  word <- values[[1]]
  embeddings_index[[word]] <- as.double(values[-1])
}

embedding_dim <- 100
embedding_matrix <- array(0, c(max_words, embedding_dim))
for (word in names(word_index)) {
  index <- word_index[[word]]
  if (index < max_words) {
    embedding_vector <- embeddings_index[[word]]
    if (!is.null(embedding_vector))
      # Words not found in the embedding index will be all zeros.
      embedding_matrix[index+1,] <- embedding_vector
  }
}

model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
summary(model)

get_layer(model, index = 1) %>% 
  set_weights(list(embedding_matrix)) %>% 
  freeze_weights()

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
save_model_weights_hdf5(model, "pre_trained_glove_model.h5")
plot(history)
```

Now compare with task-specific embedding
```{r}
model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
plot(history)
```
When training sample size = 600, the two models performed comparably. The pre-trained model has a few spikes in accuracy; the task-specific model is mode stable.


Now increase the training sample size to 1000.
Pre-trained word embeddings.
```{r}
imdb_dir <- "/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/aclImdb"
train_dir <- file.path(imdb_dir, "train")
labels <- c()
texts <- c()
for (label_type in c("neg", "pos")) {
  label <- switch(label_type, neg = 0, pos = 1)
  dir_name <- file.path(train_dir, label_type)
  for (fname in list.files(dir_name, pattern = glob2rx("*.txt"), 
                           full.names = TRUE)) {
    texts <- c(texts, readChar(fname, file.info(fname)$size))
    labels <- c(labels, label)
  }
}

library(keras)
maxlen <- 150                 
training_samples <- 1000      
validation_samples <- 10000  
max_words <- 10000            
tokenizer <- text_tokenizer(num_words = max_words) %>% 
  fit_text_tokenizer(texts)
sequences <- texts_to_sequences(tokenizer, texts)
word_index = tokenizer$word_index
cat("Found", length(word_index), "unique tokens.\n")

data <- pad_sequences(sequences, maxlen = maxlen)
labels <- as.array(labels)
cat("Shape of data tensor:", dim(data), "\n")

cat('Shape of label tensor:', dim(labels), "\n")

indices <- sample(1:nrow(data))
training_indices <- indices[1:training_samples]
validation_indices <- indices[(training_samples + 1): 
                              (training_samples + validation_samples)]
x_train <- data[training_indices,]
y_train <- labels[training_indices]
x_val <- data[validation_indices,]
y_val <- labels[validation_indices]

glove_dir = '/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/glove.6B'
lines <- readLines(file.path(glove_dir, "glove.6B.100d.txt"))
embeddings_index <- new.env(hash = TRUE, parent = emptyenv())
for (i in 1:length(lines)) {
  line <- lines[[i]]
  values <- strsplit(line, " ")[[1]]
  word <- values[[1]]
  embeddings_index[[word]] <- as.double(values[-1])
}

embedding_dim <- 100
embedding_matrix <- array(0, c(max_words, embedding_dim))
for (word in names(word_index)) {
  index <- word_index[[word]]
  if (index < max_words) {
    embedding_vector <- embeddings_index[[word]]
    if (!is.null(embedding_vector))
      # Words not found in the embedding index will be all zeros.
      embedding_matrix[index+1,] <- embedding_vector
  }
}

model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
summary(model)

get_layer(model, index = 1) %>% 
  set_weights(list(embedding_matrix)) %>% 
  freeze_weights()

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
save_model_weights_hdf5(model, "pre_trained_glove_model.h5")
plot(history)
```

Now compare with task-specific embedding
```{r}
model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
plot(history)
```
With a sample size of 1000, task-specific model is starting to show davantage. 

Now increase the training sample size to 2000.
Pre-trained word embeddings.
```{r}
imdb_dir <- "/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/aclImdb"
train_dir <- file.path(imdb_dir, "train")
labels <- c()
texts <- c()
for (label_type in c("neg", "pos")) {
  label <- switch(label_type, neg = 0, pos = 1)
  dir_name <- file.path(train_dir, label_type)
  for (fname in list.files(dir_name, pattern = glob2rx("*.txt"), 
                           full.names = TRUE)) {
    texts <- c(texts, readChar(fname, file.info(fname)$size))
    labels <- c(labels, label)
  }
}

library(keras)
maxlen <- 150                 
training_samples <- 2000     
validation_samples <- 10000  
max_words <- 10000            
tokenizer <- text_tokenizer(num_words = max_words) %>% 
  fit_text_tokenizer(texts)
sequences <- texts_to_sequences(tokenizer, texts)
word_index = tokenizer$word_index
cat("Found", length(word_index), "unique tokens.\n")

data <- pad_sequences(sequences, maxlen = maxlen)
labels <- as.array(labels)
cat("Shape of data tensor:", dim(data), "\n")

cat('Shape of label tensor:', dim(labels), "\n")

indices <- sample(1:nrow(data))
training_indices <- indices[1:training_samples]
validation_indices <- indices[(training_samples + 1): 
                              (training_samples + validation_samples)]
x_train <- data[training_indices,]
y_train <- labels[training_indices]
x_val <- data[validation_indices,]
y_val <- labels[validation_indices]

glove_dir = '/Users/Dolores/Desktop/R/MIS-64061-Advanced-Machine-Learning/glove.6B'
lines <- readLines(file.path(glove_dir, "glove.6B.100d.txt"))
embeddings_index <- new.env(hash = TRUE, parent = emptyenv())
for (i in 1:length(lines)) {
  line <- lines[[i]]
  values <- strsplit(line, " ")[[1]]
  word <- values[[1]]
  embeddings_index[[word]] <- as.double(values[-1])
}

embedding_dim <- 100
embedding_matrix <- array(0, c(max_words, embedding_dim))
for (word in names(word_index)) {
  index <- word_index[[word]]
  if (index < max_words) {
    embedding_vector <- embeddings_index[[word]]
    if (!is.null(embedding_vector))
      # Words not found in the embedding index will be all zeros.
      embedding_matrix[index+1,] <- embedding_vector
  }
}

model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
summary(model)

get_layer(model, index = 1) %>% 
  set_weights(list(embedding_matrix)) %>% 
  freeze_weights()

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
save_model_weights_hdf5(model, "pre_trained_glove_model.h5")
plot(history)
```

Now compare with task-specific embedding
```{r}
model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
plot(history)
```
At this point, the task-specific model is outperforming the pre-trained model. 

